Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-12-19 parts/def-ph.tex
refactor successive reachability in metazoan example
'saturated' GLCs + several fixes
revise TCRSIG example + justifictions of our method over [13]
minor comments of the reviewers

\todo{Gluer : explication des définitions, etc.}

@@

\subsection{Definition of the Process Hitting with $k$ classes of priorities}
\label{ssec:PH}
A PH with $k \in \sN^*$ classes of priorities (\pref{def:ph}), also simply called “PH” in the following when it is not ambiguous, gathers a finite number of concurrent \emph{processes} divided into a finite set of \emph{sorts}.
A process belongs to a unique sort and is noted $a_i$ where $a$ is the sort and $i$ the identifier of the process within the sort $a$.
Each process stands for a kind of “activity level” of its sort; a state of the PH thus corresponds to a set of processes containing exactly one process of each sort.

@@

\begin{definition}[Process Hitting with $k$ classes of priorities]
\label{def:ph}
  If $k \in \sN^*$, a \emph{Process Hitting with $k$ classes of priorities} is a triplet $\PH = (\PHs; \PHl; \PHa^{\langle k \rangle})$,
  where $\PHa^{\langle k \rangle} = (\PHa^{(1)}; \dots; \PHa^{(k)})$ is a $k$-tuple and:
  \begin{itemize}
    \item $\PHs \DEF \{a, b, \dots\}$ is the finite set of \emph{sorts},
    \item $\PHl \DEF \underset{a \in \PHs}{\times} \PHl_a$ is the finite set of \emph{states}, where $\PHl_a = \{a_0, \ldots, a_{l_a}\}$ is the finite set of \emph{processes} of sort $a \in \PHs$ and $l_a \in \sN^*$. Each process belongs to a unique sort: $\forall (a_i; b_j) \in \PHl_a \times \PHl_b, a \neq b \Rightarrow a_i \neq b_j$,
    \item $\forall n \in \llbracket 1; k \rrbracket, \PHa^{(n)} \DEF \{\PHfrappe{a_i}{b_j}{b_l} \mid (a; b) \in \PHs^2 \wedge (a_i; b_j; b_l) \in \PHl_a \times \PHl_b \times \PHl_b \wedge b_j \neq b_l \wedge a = b \Rightarrow a_i = b_j \}$ is the finite set of \emph{actions of priority $n$}.
  \end{itemize}
  We call $\PHproc \DEF \bigcup_{a \in \PHs} \PHl_a$ the set of all processes, and $\PHh \DEF \bigcup_{n \in \segm{1}{k}} \PHh^{(n)}$ the set of all actions.
  Furthermore, for all $n \in \sN^*$ and $h \in \PHh^{(n)}$, $\prio(h) \DEF n$.
\end{definition}
%
\noindent
The sort of a process $a_i$ is referred to as $\PHsort(a_i) = a$.
Given a state $s\in \PHl$, the process of sort $a \in \PHs$ present in $s$ is denoted by $\PHget{s}{a}$, that is the $a$-coordinate of the state $s$.
If $a_i \in \PHl_a$, we define the notation $a_i \in s \EQDEF \PHget{s}{a} = a_i$.
\pref{def:substate} defines the notion of sub-state on a set of sorts, that is used to focus on the interesting part of a complete state.
Note that a state is \textit{a fortiori} a sub-state.
The override of a state $s$ by a process $a_i$ (or by a set of processes)
is defined in \pref{def:statecap} as the same state in which the process of sort $a$ has been replaced by $a_i$,
which then allows to define the dynamics of a PH in \pref{def:play}.
%
\begin{definition}[Sub-states ($\PHsublize{\PHl}$)]
\label{def:substate}
  If $S \subset \PHs$ is a set of sorts, a sub-set on $S$ is an element of:
  $\PHsubl[\PHl]_S \DEF \bigtimes{a \in S} \PHl_a$.
  The set of all sub-sets is:
  $\PHsubl[\PHl] \DEF \bigcup_{S \in\powerset(\PHs)} \PHsubl[\PHl]_S$.

  \noindent
  Furthermore, if $\mysigma \in \PHsubl[\PHl]$ and $s \in \PHl$, we note:
    \[\mysigma \subseteq s \EQDEF \forall a_i \in \Proc, a_i \in \mysigma \Rightarrow a_i \in s \enspace.\]
\end{definition}
%
\begin{definition}[$\Cap : \PHl \times \PHproc \rightarrow \PHl$]
\label{def:statecap}
  Given a state $s \in \PHl$ and a process $a_i \in \PHproc$, $(s \Cap a_i)$ is the state defined by:
  $\PHget{(s \Cap a_i)}{a} = a_i \wedge \forall b \neq a, \PHget{(s \Cap a_i)}{b} = \PHget{s}{b}$.
  We also extend this definition to a set of processes given that all processes are from different sorts by the override of each process:
  $\forall ps \in \PHsubl[\PHl], s \Cap \toset{ps} = s \underset{a_i \in \toset{ps}}{\Cap} a_i$.
\end{definition}
%
\begin{definition}[Dynamics of a PH ($\PHPtrans$)]
\label{def:play}
  An action $h = \PHhit{a_i}{b_j}{b_k} \in \PHa^{(n)}$ of priority $n$ is \emph{playable} in $s \in \PHl$
  if and only if $\PHget{s}{a} = a_i$, $\PHget{s}{b} = b_j$ and $\forall m < n, \forall g \in \PHa^{(m)}, \PHhitter(g) \notin s \vee \PHtarget(g) \notin s$.
  In such a case, $(s \PHplay h)$ stands for the state resulting from the play of the action $h$ in $s$ and is defined by: $(s \PHplay h) = s \Cap b_k$.
  Moreover, we denote: $s \PHPtrans (s \PHplay h)$.

  If $s \in \PHl$,
  a \emph{scenario} $\delta$ from $s$ is a sequence of actions of $\PHh$ that can be played successively in $s$.
  The set of all scenarios from $s$ is noted $\Sce(s)$.
\end{definition}

In \pref{def:restriction}, we define the $n$-reduction of a given PH as the PH with $n$ classes of priorities in which only actions of priority lower or equal to $n$ are considered.
\begin{definition}[PH $n$-reduction]
\label{def:restriction}
  If $\PH = (\PHs; \PHl; \PHa^{\langle k \rangle})$ is a Process Hitting with $k$ classes of priorities and $n \in \segm{1}{k}$, we denote $\restriction{\PH}{n}$
  the $n$-reduction of $\PH$.
  $\restriction{\PH}{n} = (\PHs; \PHl; \PHa'^{\langle n \rangle})$ is a PH with $n$ classes of priorities with:
  $$\PHa'^{\langle n \rangle} = (\PHa^{(1)}; \dots; \PHa^{(n)})$$
  Furthermore, we denote: $\restriction{\Sce}{n}(s)$ the set of scenarios from $s$ in $\restriction{\PH}{n}$.
\end{definition}

\begin{example}
  \pref{fig:ph-livelock} gives an example of PH with $2$ classes of priorities where:
  \begin{align*}
    \PHs &= \{ a, b, c, ab \} \enspace, \\
    \PHl_a &= \{ a_0, a_1 \} \enspace, & \PHl_b &= \{ b_0, b_1 \} \enspace, \\
    \PHl_c &= \{ c_0, c_1 \} \enspace, & \PHl_{ab} &= \{ ab_{00}, ab_{01}, ab_{10}, ab_{11} \} \\ \enspace,
    \PHa^{\langle 2 \rangle} &= (\PHa^{(1)}; \PHa^{(2)})
  \end{align*}
  and the sets of actions are:
  \begin{align*}
    \PHh^{(1)} = \{ \quad
      & \PHfrappe{a_1}{ab_{00}}{ab_{10}} \enspace , \quad
      \PHfrappe{a_1}{ab_{01}}{ab_{11}} \enspace , \\
      & \PHfrappe{a_0}{ab_{10}}{ab_{00}} \enspace , \quad
      \PHfrappe{a_0}{ab_{11}}{ab_{01}} \enspace , \\
      & \PHfrappe{b_1}{ab_{00}}{ab_{01}} \enspace , \quad
      \PHfrappe{b_1}{ab_{10}}{ab_{11}} \enspace , \\
      & \PHfrappe{b_0}{ab_{01}}{ab_{00}} \enspace , \quad
      \PHfrappe{b_0}{ab_{11}}{ab_{10}}
    \quad \} \\
    \PHh^{(2)} = \{ \quad
      & \PHfrappe{a_1}{a_1}{a_0} \enspace , \quad
      \PHfrappe{b_0}{a_0}{a_1} \enspace , \\
      & \PHfrappe{b_1}{b_1}{b_0} \enspace , \quad
      \PHfrappe{a_0}{b_0}{b_1} \enspace , \\
      & \PHfrappe{ab_{11}}{c_0}{c_1}
    \quad \} \\
  \end{align*}
%   Note that: $\{ \PHhit{ab_{11}}{c_0}{c_1}, \PHhit{a_1}{a_1}{a_0}, \PHhit{a_0}{b_0}{b_1} \} \subseteq \PHh^{(2)}$.

\begin{figure}[tb]
  \centering
  \scalebox{1.2}{
  \begin{tikzpicture}
    \TSort{(0,0)}{a}{2}{l}
    \TSort{(0,4)}{b}{2}{l}
    \TSort{(7,2.5)}{c}{2}{r}

@@

    \THit{a_0}{prio}{ab_3}{.south}{ab_1}
    \THit{a_0}{prio}{ab_2}{.south}{ab_0}
    \THit{a_1}{prio}{ab_1}{.south}{ab_3}
    \THit{a_1}{prio}{ab_0}{.south}{ab_2}

    \THit{b_0}{prio}{ab_3}{.north}{ab_2}
    \THit{b_0}{prio}{ab_1}{.north}{ab_0}
    \THit{b_1}{prio}{ab_2}{.north}{ab_3}
    \THit{b_1}{prio}{ab_0}{.north}{ab_1}

    \THit{a_1}{selfhit}{a_1}{.west}{a_0}
    \THit{b_1}{selfhit}{b_1}{.west}{b_0}
    \THit{a_0.north}{bend left}{b_0}{.west}{b_1}
    \THit{b_0.south}{bend right=60}{a_0}{.west}{a_1}

@@

	\path[bounce, bend right=55]
      \TBounce{ab_0}{}{ab_2}{.west}
      \TBounce{ab_1}{}{ab_3}{.west}
	;
	\path[bounce, bend left=20]
      \TBounce{ab_3}{}{ab_1}{.south east}
      \TBounce{ab_2}{}{ab_0}{.south east}
	;
    \path[bounce, bend right=20]
      \TBounce{ab_3}{}{ab_2}{.north east}
      \TBounce{ab_1}{}{ab_0}{.north east}
    ;
    \path[bounce, bend left=30]
      \TBounce{ab_2}{}{ab_3}{.west}
      \TBounce{ab_0}{}{ab_1}{.west}
    ;
    \path[bounce, bend right]
      \TBounce{a_1}{}{a_0}{.north west}
      \TBounce{b_1}{}{b_0}{.north west}
    ;
    \path[bounce, bend left]
      \TBounce{a_0}{}{a_1}{.south west}
      \TBounce{b_0}{}{b_1}{.south west}
    ;
    \path[bounce, bend left]
      \TBounce{c_0}{}{c_1}{.south west}
    ;
    \TState{a_1, b_0, ab_2, c_0}
  \end{tikzpicture}
  }
  \caption{
  \label{fig:ph-livelock}
    An example of PH with $2$ classes of priorities.
    Sorts are represented as labelled boxes and processes as circles with their identifier on the side.
    Actions of $\PHh^{(1)}$ are represented by thick arrows and actions of $\PHh^{(2)}$ are represented by thin arrows;
    the hit part of each action in drawn in plain line and the bounce part is in dotted line.
    Greyed processes stand for a possible state $s = \PHstate{a_1, b_0, c_0, ab_{10}}$.
  }
\end{figure}

@@

\todo{Rajouter un exemple}

Cooperation between processes to make another process bounce can be expressed in PH by building a \emph{cooperative sort},
which allows to encode any kind of Boolean function \cite{PMR10-TCSB}.
\pref{fig:ph-livelock} shows an example of cooperation between processes $a_1$ and $b_1$ to make $c_0$ bounce to $c_1$:
a cooperative sort $ab$ is defined with 4 processes (one for each sub-state of the presence of processes $a_1$ and $b_1$).
For the sake of clarity, the processes of $ab$ are indexed using the sub-state they represent.
Hence, $ab_{10}$ represents the sub-state $\PHstate{a_1,b_0}$, and so on.
Each process of sort $a$ and $b$ hit $ab$ to make it bounce to the process reflecting the status of the sorts $a$ and $b$
(\eg $\PHfrappe{a_1}{ab_{00}}{ab_{10}}$ and $\PHfrappe{a_1}{ab_{01}}{ab_{11}}$).
Then, to represent the cooperation between $a_1$ and $b_1$, the process $ab_{11}$ hits $c_0$ to make it bounce to $c_1$ instead of independent hits from $a_1$ and $b_1$.

@@

\modMF{
\subsection{Justification of the priorities}
}

@@

\subsection{Restrictions}
\label{ssec:hypothesis}

@@

\pref{cr:bounded} states that the dynamics of the studied model $\PH$ contain no infinite sequence of prioritised actions.
As these actions can be considered as non-biological and therefore instantaneous, we thus prevent the existence of any Zeno-like behaviour
which would allow the play of an infinite sequence of prioritised actions in “zero time”.
As a consequence, the set $\restriction{\Sce}{1}$ is finite.

@@

Given a sort $a\in\Sigma$ and a state $s\in L$, 
we denote $\pfp_s(a)$ (\pref{def:pfp}) the processes of sort $a$ that can be present after
playing all actions of priority $1$,
which is always defined because of \pref{cr:bounded}.
\modMF{%
This definition will be useful for cooperative sorts,
in order to determine their local stable state in every configuration,
and thus which configuration each of their processes represent.
}%

@@

Our second condition (\pref{cr:compcs}) imposes that any sort in the Process
Hitting is either a well-formed component (\pref{def:component}) or a well-formed cooperative sort
(\pref{def:cs}).
The former is a sort that is hit only by actions with priority greater than $2$,
\modMF{%
and thus stands for a regular biological component.
}%
The latter is a sort that is hit only by actions with priority $1$ and which
always converges to the same process with respect to the state of the components
that have an influence on it (\pref{def:conn}).
\modMF{%
Indeed, we require that each process of a cooperative sort corresponds
to exactly one configuration of the sorts it represents.
}%
We note $\components\subset\Sigma$ the set of well-formed components,
and $\cs\subset\Sigma$ the set of well-formed cooperative sorts.

\begin{definition}[Well-formed component ($\components$)]
\label{def:component}
A sort $a \in \PHs$ is a \emph{well-formed component} if and only if:
    $\forall h \in \PHh, \PHsort(\PHtarget(h)) = a \Rightarrow \prio(h) \geq 2$ \enspace.
\end{definition}

\begin{definition}[Components influence $(\compin: \PHs\to\powerset(\components))$]
\label{def:conn}
Given a sort $a$, $\compin(a)\DEF \conn(a)\cap\components$ where
$\conn(a) \subset \PHs$ is the smallest set of sorts satisfying the following properties:
\begin{align*}
a\in\conn(a) &
\\
\forall h\in  \PHh^{(1)},
	\Sigma(\target{h})\in\conn(a) & \Longrightarrow \Sigma(\hitter{h})\in\conn(a)
\end{align*}
\end{definition}

@@

\begin{definition}[Well-formed cooperative sorts ($\cs$)]
\label{def:cs}
A sort $a \in \PHs$ is a \emph{well-formed cooperative} sort if and only if:
\begin{enumerate}[(i)]
\item $\forall h\in\PHh, \Sigma(\PHtarget(h))=a \Longrightarrow \prio(h) = 1$;
\item\label{csai} $\forall s\in\PHl, \card{\pfp_s(a)} = 1$;
\item\label{css} $\forall a_i \in \PHl_a, \exists s \in \PHl, a_i\in\pfp_s(a)$;
\item $\forall \sigma\in\PHsubl[\PHl]_{\compin(a)},
			\forall s,s'\in\PHl,
				\sigma\subseteq s\wedge \sigma\subseteq s'\Rightarrow 
							\pfp_s(a) = \pfp_{s'}(a)$\enspace.
\end{enumerate}
\end{definition}

@@

Because of \pref{def:cs}(\ref{csai}), we denote in the following: $\pfp_s(a) = a_i$.
Furthermore, because of \pref{def:cs}(\ref{css}), we denote $\csState(a_i)$ the set of sub-states represented by the process $a_i$ of any cooperative sort $a$ (\pref{def:csState}).
\begin{definition}[$\csState : \PHproc \rightarrow \powerset(\PHproc)$]
\label{def:csState}
  If $a \in \cs$ and $a_i \in \PHl_a$, 
  $$\csState(a_i) \DEF \{ \toset{ps} \mid ps\in\PHsubl[\PHl]_{\compin(a)} \wedge
  							\exists s\in\PHl, (ps\subset s\wedge \pfp_s(a)=a_i)
							\}$$
\end{definition}

@@

\subsection{Consequences of the restrictions}

>>>>

@@

\subsection{Definition of the Process Hitting with $k$ classes of priorities}
\label{ssec:PH}

A PH with $k \in \sN^*$ classes of priorities (\pref{def:ph}), also simply called “PH” in the following when it is not ambiguous, gathers a finite number of concurrent \emph{processes} divided into a finite set of \emph{sorts}.
A process belongs to a unique sort and is noted $a_i$ where $a$ is the sort and $i$ the identifier of the process within the sort $a$.
Each process stands for a kind of “activity level” of its sort; a state of the PH thus corresponds to a set of processes containing exactly one process of each sort.

@@

\begin{definition}[Process Hitting with $k$ classes of priorities]
\label{def:ph}
  If $k \in \sN^*$, a \emph{Process Hitting with $k$ classes of priorities} is a triplet $\PH = (\PHs; \PHl; \PHa^{\langle k \rangle})$,
  where $\PHa^{\langle k \rangle} = (\PHa^{(1)}; \dots; \PHa^{(k)})$ is a $k$-tuple and:
  \begin{itemize}
    \item $\PHs \DEF \{a, b, \dots, z\}$ is the finite set of \emph{sorts},
    \item $\PHl \DEF \underset{a \in \PHs}{\times} \PHl_a$ is the finite set of \emph{states}, where $\PHl_a = \{a_0, \ldots, a_{l_a}\}$ is the finite set of \emph{processes} of sort $a \in \PHs$ and $l_a \in \sN^*$. Each process belongs to a unique sort: $\forall (a_i; b_j) \in \PHl_a \times \PHl_b, a \neq b \Rightarrow a_i \neq b_j$,
    \item $\forall n \in \llbracket 1; k \rrbracket, \PHa^{(n)} \DEF \{\PHfrappe{a_i}{b_j}{b_l} \mid (a; b) \in \PHs^2 \wedge (a_i; b_j; b_l) \in \PHl_a \times \PHl_b \times \PHl_b \wedge b_j \neq b_l \wedge a = b \Rightarrow a_i = b_j \}$ is the finite set of \emph{actions of priority $n$}.
  \end{itemize}
  We call $\PHproc \DEF \bigcup_{a \in \PHs} \PHl_a$ the set of all processes, and $\PHh \DEF \bigcup_{n \in \segm{1}{k}} \PHh^{(n)}$ the set of all actions.
  Furthermore, for all $h \in \PHh$,
  %for all $n \in \sN^*$ and $h \in \PHh^{(n)}$,
  $\prio(h) \DEF \min\{ n \in \segm{1}{k} \mid h \in \PHh^{(n)} \}$.
\end{definition}
%
\noindent
The sort of a process $a_i$ is referred to as $\PHsort(a_i) = a$.
Given a state $s\in \PHl$, the process of sort $a \in \PHs$ present in $s$ is denoted by $\PHget{s}{a}$, that is the $a$-coordinate of the state $s$.
If $a_i \in \PHl_a$, we define the notation $a_i \in s \EQDEF \PHget{s}{a} = a_i$.
\pref{def:substate} defines the notion of sub-state on a set of sorts, that is used to focus on the interesting part of a complete state.
Note that a state is \textit{a fortiori} a sub-state.
The override of a state $s$ by a process $a_i$ (or by a set of processes)
is defined in \pref{def:statecap} as the same state in which the process of sort $a$ has been replaced by $a_i$,
which then allows to define the dynamics of a PH in \pref{def:play}.
%
\begin{definition}[Sub-states ($\PHsublize{\PHl}$)]
\label{def:substate}
  If $S \subseteq \PHs$ is a set of sorts, a sub-state on $S$ is an element of:
  $\PHsubl[\PHl]_S \DEF \bigtimes{a \in S} \PHl_a$.
  The set of all sub-states is:
  $\PHsubl[\PHl] \DEF \bigcup_{S \in\powerset(\PHs)} \PHsubl[\PHl]_S$.

  \noindent
  Furthermore, if $\mysigma \in \PHsubl[\PHl]$ and $s \in \PHl$, we note:
    \[\mysigma \subseteq s \EQDEF \forall a_i \in \Proc, a_i \in \mysigma \Rightarrow a_i \in s \enspace.\]
\end{definition}
%
\begin{definition}[$\Cap : \PHl \times \PHproc \rightarrow \PHl$]
\label{def:statecap}
  Given a state $s \in \PHl$ and a process $a_i \in \PHproc$, $(s \Cap a_i)$ is the state defined by:
  $\PHget{(s \Cap a_i)}{a} = a_i \wedge \forall b \neq a, \PHget{(s \Cap a_i)}{b} = \PHget{s}{b}$.
  Under the condition that all processes are from different sorts,
  we can also extend this definition to a set of processes
  by the override of each process independently:
  $\forall ps \in \PHsubl[\PHl], s \Cap \toset{ps} = s \underset{a_i \in \toset{ps}}{\Cap} a_i$.
\end{definition}
The notation $\toset{ps}$ represents the set of all elements in $ps$
(which is not a set but a tuple)
as defined on page~\pageref{notations}.
%
\begin{definition}[Dynamics of a PH ($\PHPtrans$)]
\label{def:play}
  An action $h = \PHhit{a_i}{b_j}{b_k} \in \PHa^{(n)}$ of priority $n$ is \emph{playable} in $s \in \PHl$
  if and only if $\PHget{s}{a} = a_i$, $\PHget{s}{b} = b_j$ and $\forall m < n, \forall g \in \PHa^{(m)}, \PHhitter(g) \notin s \vee \PHtarget(g) \notin s$.
  In such a case, $(s \PHplay h)$ stands for the state resulting from the play of the action $h$ in $s$ and is defined by: $(s \PHplay h) = s \Cap b_k$.
  Moreover, we denote: $s \PHPtrans (s \PHplay h)$.

  If $s \in \PHl$,
  a \emph{scenario} $\delta$ from $s$ is a (possibly empty) sequence of actions of $\PHh$
  that can be played successively in $s$.
  The set of all scenarios from $s$ is noted $\Sce(s)$.
\end{definition}

\modMF{
In \pref{def:restriction}, we define the $1$-reduction of a given PH as the model
in which only actions of highest priority lower are considered.
\begin{definition}[PH $1$-reduction]
\label{def:restriction}
  If $\PH = (\PHs; \PHl; \PHa^{\langle k \rangle})$ is a Process Hitting with $k$ classes of priorities, we denote by $\restriction{\PH}{1}$ the $1$-reduction of $\PH$:
  $\restriction{\PH}{1} = (\PHs; \PHl; \PHa'^{\langle 1 \rangle})$
  with $\PHa'^{\langle 1 \rangle} = (\PHa^{(1)})$.
  Furthermore, we denote by $\restriction{\Sce}{1}(s)$ the set of scenarios from $s$ in $\restriction{\PH}{1}$.
\end{definition}
}

\begin{example}
  \pref{fig:ph-livelock} gives an example of PH with $2$ classes of priorities where:
  \begin{align*}
    \PHs &= \{ a, b, c, ab \} \enspace,
      & \PHa^{\langle 2 \rangle} &= (\PHa^{(1)}; \PHa^{(2)}) \enspace, \\
    \PHl_a &= \{ a_0, a_1 \} \enspace,
      & \PHl_b &= \{ b_0, b_1 \} \enspace, \\
    \PHl_c &= \{ c_0, c_1 \} \enspace,
      & \PHl_{ab} &= \{ ab_{00}, ab_{01}, ab_{10}, ab_{11} \}
  \end{align*}
  and the sets of actions are:
  \begin{align*}
    \PHh^{(1)} = \{ \quad
      & \PHfrappe{a_1}{ab_{00}}{ab_{10}} \enspace , \quad
      \PHfrappe{a_1}{ab_{01}}{ab_{11}} \enspace , \\
      & \PHfrappe{a_0}{ab_{10}}{ab_{00}} \enspace , \quad
      \PHfrappe{a_0}{ab_{11}}{ab_{01}} \enspace , \\
      & \PHfrappe{b_1}{ab_{00}}{ab_{01}} \enspace , \quad
      \PHfrappe{b_1}{ab_{10}}{ab_{11}} \enspace , \\
      & \PHfrappe{b_0}{ab_{01}}{ab_{00}} \enspace , \quad
      \PHfrappe{b_0}{ab_{11}}{ab_{10}}
    \quad \} \\
    \PHh^{(2)} = \{ \quad
      & \PHfrappe{a_1}{a_1}{a_0} \enspace , \quad
      \PHfrappe{b_0}{a_0}{a_1} \enspace , \\
      & \PHfrappe{b_1}{b_1}{b_0} \enspace , \quad
      \PHfrappe{a_0}{b_0}{b_1} \enspace , \\
      & \PHfrappe{ab_{11}}{c_0}{c_1}
    \quad \} \\
  \end{align*}
%   Note that: $\{ \PHhit{ab_{11}}{c_0}{c_1}, \PHhit{a_1}{a_1}{a_0}, \PHhit{a_0}{b_0}{b_1} \} \subseteq \PHh^{(2)}$.

\begin{figure}[tb]
  \centering
  \scalebox{1.3}{
  \begin{tikzpicture}
    \path[use as bounding box] (-.2,-.5) rectangle (7.2,5.7);

    \TSort{(0,0)}{a}{2}{l}
    \TSort{(0,4)}{b}{2}{l}
    \TSort{(7,2.5)}{c}{2}{r}

@@

    \THit{a_0.-20}{bend right,in=270,prio}{ab_3}{.south}{ab_1}
    \THit{a_0.20}{bend right,in=250,prio}{ab_2}{.south}{ab_0}
    \THit{a_1.-20}{bend right,in=240,prio}{ab_1}{.south}{ab_3}
    \THit{a_1.20}{bend right,in=220,prio}{ab_0}{.south}{ab_2}

    \THit{b_0}{bend right,in=70,out=40,prio}{ab_3}{.north}{ab_2}
    \THit{b_0.south west}{bend right,in=140,out=40,prio}{ab_1}{.north}{ab_0}
    \THit{b_1}{bend right,in=90,out=40,prio}{ab_2}{.north}{ab_3}
    \THit{b_1}{bend right,in=170,out=10,prio}{ab_0}{.north}{ab_1}

    \THit{a_1}{selfhit}{a_1}{.west}{a_0}
    \THit{b_1}{selfhit}{b_1}{.west}{b_0}
    \THit{a_0.110}{bend left}{b_0}{.210}{b_1}
    \THit{b_0.-10}{bend right=60}{a_0}{.west}{a_1}

@@

	\path[bounce, bend right=90] %55]
      \TBounce{ab_0}{in=-110}{ab_2}{.south west}
      \TBounce{ab_1}{in=-110}{ab_3}{.south west}
	;
	\path[bounce, bend left=25]
      \TBounce{ab_3}{}{ab_1}{.south east}
      \TBounce{ab_2}{}{ab_0}{.south east}
	;
    \path[bounce, bend right=30]
      \TBounce{ab_3}{}{ab_2}{.30}
      \TBounce{ab_1}{}{ab_0}{.30}
    ;
    \path[bounce, bend left=40]
      \TBounce{ab_2}{}{ab_3}{.150}
      \TBounce{ab_0}{}{ab_1}{.150}
    ;
    \path[bounce, bend right=60]
      \TBounce{a_1}{}{a_0}{.north west}
      \TBounce{b_1}{}{b_0}{.north west}
    ;
    \path[bounce, bend left=60]
      \TBounce{a_0}{}{a_1}{.south west}
      \TBounce{b_0}{}{b_1}{.south west}
    ;
    \path[bounce, bend left]
      \TBounce{c_0}{}{c_1}{.south west}
    ;
    \TState{a_1, b_0, ab_2, c_0}
  \end{tikzpicture}
  }
  \caption{%
  \label{fig:ph-livelock}%
    An example of PH with $2$ classes of priorities.
    This model represents the interaction of two exclusive components $a$ and $b$,
    that cannot be active simultaneously, and that degrade over time.
    Furthermore, the cooperative sort $ab$ allows to encode a conjunction between
    the processes $a_1$ and $b_1$ in order to activate the third component $c$.
    Sorts are represented by labelled boxes and processes by circles with their identifier on the side.
    Actions of $\PHh^{(1)}$ are represented by thick arrows and actions of $\PHh^{(2)}$ are represented by thin arrows;
    the hit part of each action is drawn in plain line and the bounce part is in dotted line.
    Greyed processes stand for the following possible state: $\PHstate{a_1, b_0, c_0, ab_{10}}$.
  }
\end{figure}

@@

Cooperation between processes to make another process bounce can be expressed in PH by building a \emph{cooperative sort},
which allows to encode any kind of Boolean function \cite{PMR10-TCSB}.
\pref{fig:ph-livelock} shows an example of cooperation between processes $a_1$ and $b_1$
in order to allow $c_0$ to bounce to $c_1$.
In this model, a cooperative sort $ab$ is defined with 4 processes
in order to represent each possible configuration
of the presence of processes $a_1$ and $b_1$
(“only $a_1$ is active”, “only $b_1$ is active”, “both are active” and “both are not active”).
For the sake of clarity, the processes of $ab$ are indexed using the sub-state they represent.
Hence, $ab_{10}$ represents the sub-state $\PHstate{a_1,b_0}$, and so on.
Each process of sort $a$ and $b$ hit $ab$ to make it bounce to the process reflecting the status of the sorts $a$ and $b$
(\eg $\PHfrappe{a_1}{ab_{00}}{ab_{10}}$ and $\PHfrappe{a_1}{ab_{01}}{ab_{11}}$).
Then, to represent the cooperation between $a_1$ and $b_1$, the process $ab_{11}$ hits $c_0$ to make it bounce to $c_1$.
Thus, this method based on a cooperative sort
replaces independent actions, \eg $\PHhit{a_1}{c_0}{c_1}$ and $\PHhit{b_1}{c_0}{c_1}$,
that would not be sufficient to model the conjunction
between $a_1$ and $b_1$.

@@

\modMF{
\subsection{Justification of the priorities}
}
\label{ssec:priorities}

@@

\subsection{Conditions}
\label{ssec:hypothesis}

@@

\pref{cr:bounded} states that the dynamics of the studied model $\PH$ contain no infinite sequence of prioritised actions.
As these actions can be considered as non-biological and therefore instantaneous, we thus prevent the existence of any Zeno-like behaviour
which would allow the play of an infinite sequence of prioritised actions in “zero time”.
As a consequence, the set $\restriction{\Sce}{1}$,
\modMF{
that is, the set of scenarios playable only with actions of $\PHh^{(1)}$,
(see \pref{def:restriction}),
}
is finite.

@@

Given a sort $a\in\Sigma$ and a state $s\in \PHl$, 
we denote by $\pfp_s(a)$ (\pref{def:pfp}) the processes of sort $a$ that can be present after
playing all actions of priority $1$ in $s$,
which is always defined because of \pref{cr:bounded}.
\modMF{%
This definition will be useful for cooperative sorts,
in order to determine their local stable state in every configuration,
and thus which configuration each of their processes represent.
}%

@@

Our second condition (\pref{cr:compcs}) imposes that any sort in the Process
Hitting is either a well-formed component (\pref{def:component}) or a well-formed cooperative sort
(\pref{def:cs}).
The former is a sort that is hit only by actions with priority greater than or equal to $2$,
\modMF{%
and thus stands for a regular biological component.
}%
The latter is a sort that is hit only by actions with priority $1$ and which
always converges to the same process with respect to the state of the components
that have an influence on it (\pref{def:conn}).
\modMF{%
Indeed, we require that each process of a cooperative sort corresponds
to exactly one configuration of the sorts it represents.
}%
% We denote by $\components\subseteq\Sigma$ the set of well-formed components,
% and $\cs\subseteq\Sigma$ the set of well-formed cooperative sorts.
\modMF{
We note that some non-determinism may be involved during the update of
cooperative sorts.
However, this non-determinism is only confined to the choice of the order of the
actions played or the update of one cooperative sort before another.
Thus, there is no non-determinism in the focal process of each cooperative sort,
that is, from the point of view of lower-priority actions.
}

\begin{definition}[Well-formed component ($\components$)]
\label{def:component}
  A sort $a \in \PHs$ is a \emph{well-formed component} if and only if:
    $\forall h \in \PHh, \PHsort(\PHtarget(h)) = a \Rightarrow \prio(h) \geq 2$.
  We denote by $\components$ the set of all well-formed components.
\end{definition}

\begin{definition}[Components influence $(\compin: \PHs\to\powerset(\components))$]
\label{def:conn}
  Given a sort $a$, $\compin(a)\DEF \conn(a)\cap\components$ where
  $\conn(a) \subseteq \PHs$ is the smallest set of sorts satisfying the following properties:
  \begin{align*}
    a\in\conn(a) &
    \\
    \forall h\in  \PHh^{(1)},
      \Sigma(\target{h})\in\conn(a) & \Longrightarrow \Sigma(\hitter{h})\in\conn(a)
  \end{align*}
\end{definition}

@@

\begin{definition}[Well-formed cooperative sorts ($\cs$)]
\label{def:cs}
  A sort $a \in \PHs$ is a \emph{well-formed cooperative} sort if and only if:
  \begin{enumerate}[(i)]
    \item $\forall h\in\PHh, \Sigma(\PHtarget(h))=a \Longrightarrow \prio(h) = 1$;
    \item\label{csai} $\forall s\in\PHl, \card{\pfp_s(a)} = 1$;
    \item\label{css} $\forall a_i \in \PHl_a, \exists s \in \PHl, a_i\in\pfp_s(a)$;
    \item $\forall \sigma\in\PHsubl[\PHl]_{\compin(a)},
      \forall s,s'\in\PHl,
      \sigma\subseteq s\wedge \sigma\subseteq s'\Rightarrow 
      \pfp_s(a) = \pfp_{s'}(a)$\enspace.
  \end{enumerate}
  We denote by $\cs$ the set of all well-formed cooperative sorts.
\end{definition}

@@

Because of \pref{def:cs}(\ref{csai}), we denote in the following: $\pfp_s(a) = a_i$.
Furthermore, because of \pref{def:cs}(\ref{css}), we denote by $\csState(a_i)$ the set of sub-states modelled by the process $a_i$ of any cooperative sort $a$ (\pref{def:csState}),
that is, the sub-states predecessors for which the focal process of $a$ is $a_i$.
\begin{definition}[$\csState : \PHproc \rightarrow \powerset(\PHproc)$]
\label{def:csState}
  If $a \in \cs$ and $a_i \in \PHl_a$, 
  $$\csState(a_i) \DEF \{ \toset{ps} \mid ps\in\PHsubl[\PHl]_{\compin(a)} \wedge
  							\exists s\in\PHl, (ps\subseteq s\wedge \pfp_s(a)=a_i)
							\}$$
\end{definition}

@@

\subsection{Consequences of the Conditions}
